Machine Learning for Predictive Auto-Tuning with Boosted Regression Trees

### 2012-03-13 paper/final.tex
adding acknowledgement section
final touch-up

% UPDATE ABSTRCT.TXT IF YOU CHANGE THIS!!! (abstract.txt)
The rapidly evolving landscape of multicore architectures makes
the construction of efficient libraries a daunting task.
A family of methods known collectively as ``auto-tuning'' has emerged to
address this challenge.  Two major approaches to auto-tuning are
{\em empirical} and {\em model-based}:
empirical auto-tuning is a generic but slow approach that works by measuring runtimes of
candidate implementations, model-based auto-tuning predicts those runtimes using simplified abstractions designed by hand.
We show that machine learning methods for non-linear regression can be used
to estimate timing models from data, capturing the best of both approaches.
A statistically-derived model offers the speed of a model-based approach,
with the generality and simplicity of empirical auto-tuning.
We validate our approach using the filterbank
correlation kernel described in \citet{pinto+cox:2011gcg}, where we find that
0.1 seconds of hill climbing on the regression model (``{\em
predictive} auto-tuning'') can achieve an average of 95\% of the
speed-up brought by minutes of empirical auto-tuning.  Our approach is not
specific to filterbank correlation, nor even to GPU kernel auto-tuning, and can
be applied to almost any templated-code optimization problem, spanning a wide
variety of problem types, kernel types, and platforms.
% UPDATE ABSTRCT.TXT IF YOU CHANGE THIS!!! (abstract.txt)

@@

\cite{pinto+cox:2011gcg} recommends as a reference implementation:
block\_w = 8,
block\_h = 8,
n\_filter\_rows = 1,
n\_output\_4s = all,
spill = $False$,
imul\_fast = $True$,
pad\_sha-red = $True$, %%% XXX: hypenation by hand
use\_text1d = $True$,
maxrreg = $\infty$,
and fast\_math = $False$.
This reference implementation was chosen manually based on good performance
across a range of platforms from older-generation cards such as the 8600GT all
the way to current-generation flagship cards such as the GTX 580 and C2070.
Given that parameters were hand-chosen for the reference kernel, no claims are
made as to the optimality nor universality of this reference (indeed, different
programmers would undoubtedly arrive at different results).  We use
this kernel configuration as a reasonable indicator of typical performance
made possible by ad hoc experimentation with parameters.

@@

% backtracking
In some cases, predictive auto-tuning yields an invalid implementation.  We
dealt with this scenario by backtracking through the various best-estimates
found during the hill-climbing search.  Some invalid kernels can only be
discovered after compiling code and attempting to run the compiled code, so in
these cases predictive auto-tuning took up to 3 seconds.  Even in these cases
the predictive model is much faster than empirical auto-tuning, because the
first kernel that actually runs is a fast one.

@@

%Our

>>>>

% UPDATE ABSTRCT.TXT IF YOU CHANGE THIS!!! (abstract.txt)
The rapidly evolving landscape of multicore architectures makes
the construction of efficient libraries a daunting task.
A family of methods known collectively as ``auto-tuning'' has emerged to
address this challenge.  Two major approaches to auto-tuning are
{\em empirical} and {\em model-based}:
empirical auto-tuning is a generic but slow approach that works by measuring runtimes of
candidate implementations, model-based auto-tuning predicts those runtimes using simplified abstractions designed by hand.
We show that machine learning methods for non-linear regression can be used
to estimate timing models from data, capturing the best of both approaches.
A statistically-derived model offers the speed of a model-based approach,
with the generality and simplicity of empirical auto-tuning.
We validate our approach using the filterbank
correlation kernel described in \citet{pinto+cox:2011gcg}, where we find that
0.1 seconds of hill climbing on the regression model (``{\em
predictive} auto-tuning'') can achieve almost the same
speed-up as is brought by minutes of empirical auto-tuning.  Our approach is not
specific to filterbank correlation, nor even to GPU kernel auto-tuning, and can
be applied to almost any templated-code optimization problem, spanning a wide
variety of problem types, kernel types, and platforms.
% UPDATE ABSTRCT.TXT IF YOU CHANGE THIS!!! (abstract.txt)

@@

\cite{pinto+cox:2011gcg} recommend as a reference implementation:
block\_w = 8,
block\_h = 8,
n\_filter\_rows = 1,
n\_output\_4s = all,
spill = $False$,
imul\_fast = $True$,
pad\_sha-red = $True$, %%% XXX: hypenation by hand
use\_text1d = $True$,
maxrreg = $\infty$,
and fast\_math = $False$.
This reference implementation was chosen manually based on good performance
across a range of platforms from older-generation cards such as the 8600GT all
the way to current-generation flagship cards such as the GTX 580 and C2070.
Given that parameters were hand-chosen for the reference kernel, no claims are
made as to the optimality nor universality of this reference (indeed, different
programmers would undoubtedly arrive at different results).  We use
this kernel configuration as a reasonable indicator of typical performance
made possible by ad hoc experimentation with parameters.

@@

% backtracking
In some cases, predictive auto-tuning yields an invalid implementation.  We
dealt with this scenario by backtracking through the various best-estimates
found during the hill-climbing search.  Some invalid kernels can only be
discovered after compiling code and attempting to run the compiled code, so in
these cases predictive auto-tuning took up to 3 seconds.  Even in these cases
the predictive model is much faster than empirical auto-tuning because the
first kernel that runs successfully is still typically a fast one.

@@

%Our
\section{Acknowledgments}

This work was funded by the Rowland Institute of Harvard, the NVIDIA
Graduate Fellowship, and the National Science Foundation (IIS 0963668).
Hardware support was generously provided by the NVIDIA Corporation.
